Panel Seminars

### 2025-04-17 POP_10_Signatures_of_Selection.Rmd
publishing sigs of selection page

---
title: "A MarineOmics overview of methods to detect signatures of selection"
author: "Joanna Griffiths, Danielle Davenport, Madison Armstrong, Camille Rumberger"
date: "2025-02-24"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: true
bibliography: "common-bib_01.bib"
---

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 1: Example of an Fst-based analysis using [OutFLANK](https://rstudio-pubs-static.s3.amazonaws.com/305384_9aee1c1046394fb9bd8e449453d72847.html). In this dataset, 1000 individuals were collected from across 39 populations  spanning an environmental gradient. Each dot is a SNP and its location in the genome plotted on the X-axis. The Y-axis is the Fst for each SNP. Higher Fst values are indicative of larger genetic differentiation and candidate SNPs under selection are highlighted in pink. [Source](https://htmlpreview.github.io/?https://github.com/whitlock/OutFLANK/blob/master/inst/doc/OutFLANKAnalysis.html)"}
knitr::include_graphics("images/Fig1.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 2: Example of a PCA-based analysis using [pcadapt](https://bcm-uga.github.io/pcadapt/articles/pcadapt.html). In this dataset, a total of 150 individuals from three different populations were genotyped at 1,500 markers. Population structure was identified for three discrete groups of individuals, matching population assignment expectations (A). SNPs that are significantly associated with PC1 and PC2 are given higher weightings and are used as input for a Manhattan plot, where the SNP location in the genome is plotted on the X-axis and the weightings (or -log pvalue) are plotted on the Y-axis. SNPs that have higher weightings will have higher -log pvalues and are candidate SNPs under selection among populations (B). [Source](https://bcm-uga.github.io/pcadapt/articles/pcadapt.html)"}
knitr::include_graphics("images/Fig2.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 3: Example of a RDA plot. The blue vectors are the environmental predictors. Individuals are color coded by location and plotted in the ordination space based on their relationship with the axes (which are linear combinations of the predictor variables). SNPs are plotted gray in the center of the plot. Note that individual genotypes from wet and temperate British Columbia (green) are positively related to high annual precipitation (AP) and low temperature seasonality (sdT). Based on RDA results, SNPs are strongly correlated with these two precipitation values and these candidate SNPs are plotted along the AP vector out from the center of the plot. [Source](https://popgen.nescent.org/2018-03-27_RDA_GEA.html)"}
knitr::include_graphics("images/Fig3.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 4. In a study by [(@nielsen2021distinct)](#refs), GF modeling was used to infer genomic variation–environment relationships. Models were fitted to the allele frequencies of putatively neutral and outlier SNP datasets (detected with other methods) for three rocky shore species and environmental data from WorldClim and principal coordinates of neighbor matrices (PCNMs) were used as predictors in the model to account for spatial autocorrelation and unmeasured environmental variation. The figure shows cumulative importance curves based on the turnover functions for each environmental variable for the urchin (left) where the curve shape indicates the rate of change in allele frequencies along the environmental gradient, and the maximum height indicates the total turnover in allele frequencies. The authors then used the turnover functions from the GF models to transform the environmental variables into a common biological importance scale (i.e [(@fitzpatrick2015ecological)](#refs) (left). We can see neutral (a, c, e) or outlier (b, d, f) loci where the coloured map cells indicate similar allele frequencies and show that the genomic turnover patterns of putative neutral and outlier loci generally followed the biogeographical breaks of the marine environment in the region."}
knitr::include_graphics("images/Fig4.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 5:  An example of a SNP where a lower allele frequency was found to be significantly associated with warmer minimum winter temperatures using Bayenv. This SNP was characterized by a high Bayes factors and extreme allele frequencies. (A) Allele frequencies and standardized minimum winter temperatures of the SNP, rs6001912 (B) shows the geographical distribution of SNP rs6001912 [(@fitzpatrick2021experimental)](#refs)"}
knitr::include_graphics("images/Fig5.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 6: In a study on red abalone population genomics by [(@griffiths2025seascape)](#refs), lfmm2 modeling was used to infer genomic variation–environment relationships with maximum sea surface temperature (SST) (A), Coastal Upwelling Transport Index (CUTI) (B), Biologically Effective Upwelling Transport Index (BEUTI) (C) and range BEUTI (D). The horizontal line on the manhattan plots represents the significance threshold after Bonferroni correction of 0.05 (red). Very few SNPs passed the significance threshold and they are spread across multiple scaffolds, suggesting there is weak evidence for local adaptation to these environmental gradients."}
knitr::include_graphics("images/Fig6.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 7 (adapted from Figure 1 from [(@dauphin2023re)](#refs)): The green boundary represents the species regional range limit and the colors reflect topography. There are three main sampling design strategies that can be used in landscape genetic studies: (A) a sampling design that doesn’t rely on environmental or genetic data, (B) a sampling design that considers environmental data only, and (C) a sampling design that takes into account genetic and environmental information (Circles in box C represent sampling locations, with shadings referring to a priori known neutral genetic differences)."}
knitr::include_graphics("images/Fig7.jpg")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 8: Example of GWAS implemented using a linear regression in the program PLINK. 237 coral samples were collected across 12 reefs from the Great Barrier reef. Low-coverage WGS was performed and bleaching susceptibility was measured in each individual. Each dot is a SNP and it is plotted by its location in the genome on the X-axis. The Y-axis is the -log p value for the linear regression of each SNP and its association with bleaching susceptibility. Larger values demonstrate stronger associations. The authors identified two genes with significant loci that were associated with bleaching susceptibility. Source: [(@fuller2020population)](#refs)"}
knitr::include_graphics("images/Fig8.png")

>>>>

---
title: "A MarineOmics overview of methods to detect signatures of selection"
author: "Joanna Griffiths, Danielle Davenport, Madison Armstrong, Camille Rumberger"
date: "2025-02-24"
output:
  html_document:
    toc: yes
    toc_float:
      collapsed: true
bibliography: "POP_10_Signatures_Selection/common-bib_01.bib"
---

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 1: Example of an Fst-based analysis using [OutFLANK](https://rstudio-pubs-static.s3.amazonaws.com/305384_9aee1c1046394fb9bd8e449453d72847.html). In this dataset, 1000 individuals were collected from across 39 populations  spanning an environmental gradient. Each dot is a SNP and its location in the genome plotted on the X-axis. The Y-axis is the Fst for each SNP. Higher Fst values are indicative of larger genetic differentiation and candidate SNPs under selection are highlighted in pink. [Source](https://htmlpreview.github.io/?https://github.com/whitlock/OutFLANK/blob/master/inst/doc/OutFLANKAnalysis.html)"}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig1.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 2: Example of a PCA-based analysis using [pcadapt](https://bcm-uga.github.io/pcadapt/articles/pcadapt.html). In this dataset, a total of 150 individuals from three different populations were genotyped at 1,500 markers. Population structure was identified for three discrete groups of individuals, matching population assignment expectations (A). SNPs that are significantly associated with PC1 and PC2 are given higher weightings and are used as input for a Manhattan plot, where the SNP location in the genome is plotted on the X-axis and the weightings (or -log pvalue) are plotted on the Y-axis. SNPs that have higher weightings will have higher -log pvalues and are candidate SNPs under selection among populations (B). [Source](https://bcm-uga.github.io/pcadapt/articles/pcadapt.html)"}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig2.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 3: Example of a RDA plot. The blue vectors are the environmental predictors. Individuals are color coded by location and plotted in the ordination space based on their relationship with the axes (which are linear combinations of the predictor variables). SNPs are plotted gray in the center of the plot. Note that individual genotypes from wet and temperate British Columbia (green) are positively related to high annual precipitation (AP) and low temperature seasonality (sdT). Based on RDA results, SNPs are strongly correlated with these two precipitation values and these candidate SNPs are plotted along the AP vector out from the center of the plot. [Source](https://popgen.nescent.org/2018-03-27_RDA_GEA.html)"}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig3.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 4. In a study by [(@nielsen2021distinct)](#refs), GF modeling was used to infer genomic variation–environment relationships. Models were fitted to the allele frequencies of putatively neutral and outlier SNP datasets (detected with other methods) for three rocky shore species and environmental data from WorldClim and principal coordinates of neighbor matrices (PCNMs) were used as predictors in the model to account for spatial autocorrelation and unmeasured environmental variation. The figure shows cumulative importance curves based on the turnover functions for each environmental variable for the urchin (left) where the curve shape indicates the rate of change in allele frequencies along the environmental gradient, and the maximum height indicates the total turnover in allele frequencies. The authors then used the turnover functions from the GF models to transform the environmental variables into a common biological importance scale (i.e [(@fitzpatrick2015ecological)](#refs) (left). We can see neutral (a, c, e) or outlier (b, d, f) loci where the coloured map cells indicate similar allele frequencies and show that the genomic turnover patterns of putative neutral and outlier loci generally followed the biogeographical breaks of the marine environment in the region."}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig4.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 5:  An example of a SNP where a lower allele frequency was found to be significantly associated with warmer minimum winter temperatures using Bayenv. This SNP was characterized by a high Bayes factors and extreme allele frequencies. (A) Allele frequencies and standardized minimum winter temperatures of the SNP, rs6001912 (B) shows the geographical distribution of SNP rs6001912 [(@fitzpatrick2021experimental)](#refs)"}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig5.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 6: In a study on red abalone population genomics by [(@griffiths2025seascape)](#refs), lfmm2 modeling was used to infer genomic variation–environment relationships with maximum sea surface temperature (SST) (A), Coastal Upwelling Transport Index (CUTI) (B), Biologically Effective Upwelling Transport Index (BEUTI) (C) and range BEUTI (D). The horizontal line on the manhattan plots represents the significance threshold after Bonferroni correction of 0.05 (red). Very few SNPs passed the significance threshold and they are spread across multiple scaffolds, suggesting there is weak evidence for local adaptation to these environmental gradients."}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig6.png")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 7 (adapted from Figure 1 from [(@dauphin2023re)](#refs)): The green boundary represents the species regional range limit and the colors reflect topography. There are three main sampling design strategies that can be used in landscape genetic studies: (A) a sampling design that doesn’t rely on environmental or genetic data, (B) a sampling design that considers environmental data only, and (C) a sampling design that takes into account genetic and environmental information (Circles in box C represent sampling locations, with shadings referring to a priori known neutral genetic differences)."}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig7.jpg")

@@

```{r, echo=FALSE, out.width="75%", fig.align="center", fig.cap="left-aligned", fig.cap="Figure 8: Example of GWAS implemented using a linear regression in the program PLINK. 237 coral samples were collected across 12 reefs from the Great Barrier reef. Low-coverage WGS was performed and bleaching susceptibility was measured in each individual. Each dot is a SNP and it is plotted by its location in the genome on the X-axis. The Y-axis is the -log p value for the linear regression of each SNP and its association with bleaching susceptibility. Larger values demonstrate stronger associations. The authors identified two genes with significant loci that were associated with bleaching susceptibility. Source: [(@fuller2020population)](#refs)"}
knitr::include_graphics("POP_10_Signatures_Selection/images/Fig8.png")

### 2025-04-17 index.Rmd
publishing sigs of selection page

* [Choosing a population genomics sequencing approach](POP_01_choosing_population_genetics.html)  
* [Whole Genome Resequencing (WGS)](POP_04_WGS_intro.html)  
  * [Low coverage WGS](https://github.com/therkildsen-lab/genomic-data-analysis/blob/master/lcwgs_data_analysis.md)
* [Reduced Representation Sequencing](RADseq.html) 
* [Poolseq](POP_03_poolseq.html)
* [Redundancy Analysis (RDA) Trait Prediction](RDAtraitPredictionTutorial.html)
* [PCA](POP_08_PCA.html)

>>>>

* [Choosing a population genomics sequencing approach](POP_01_choosing_population_genetics.html)  
* [Whole Genome Resequencing (WGS)](POP_04_WGS_intro.html)  
  * [Low coverage WGS](https://github.com/therkildsen-lab/genomic-data-analysis/blob/master/lcwgs_data_analysis.md)
* [Reduced Representation Sequencing](RADseq.html) 
* [Poolseq](POP_03_poolseq.html)
* [Redundancy Analysis (RDA) Trait Prediction](RDAtraitPredictionTutorial.html)
* [PCA](POP_08_PCA.html)
* [Ancient & Degraded DNA](POP_09_aDNA.html)
* [Methods to detect signatures of selection](POP_10_Signatures_of_Selection.html)
